### 2023-05-18 scripts/project_template.Rmd
grab input date columns for relocation

---
title: "Barneby Lives usage example"
author: "steppe"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidyverse)
library(sf)
source('functions.R')
```

```{r dummy data}
dummy_pts <- data.frame(
  collectionNo = paste0(rep('RCB', 1000), seq(1:1000)),
  day = c(sample(1:30, size = 1000, replace = T)),
  month = c(sample(5:10, size = 1000, replace = T)),
  year = 2023,
  datum = sample(c('nad83', 'nad27', 'wgs84'), size = 1000, replace = T),
  longitude = runif(1000, min = -120, max = -100), 
  latitude = runif(1000, min = 35, max = 48)
) %>% 
  mutate(col_date = paste(month, day, year, sep = "/"))

```

```{r coordinate parsing and make object spatial in wgs84}

out <- dms2dd(dummy_pts)
out <- coords2sf(out)

```

```{r date parsing}

out <- date_parser(out, coll_date = 'col_date')

```

```{r Acquire geodata}

places <- sf::st_read('../geodata/places/places.shp', quiet = T)

out1 <- political_grabber(dummy_pts, 'collectionNo')
out2 <- place_grabber(dummy_pts)
out3 <- physical_grabber(dummy_pts)

```

```{r gather taxonomic data}

```