### 2023-05-23 scripts/project_template.Rmd
m

---
title: "Barneby Lives usage example"
author: "steppe"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidyverse)
library(sf)
source('functions.R')
```

```{r dummy data}
dummy_pts <- data.frame(
  collectionNo = paste0(rep('RCB', 1000), seq(1:1000)),
  day = c(sample(1:30, size = 1000, replace = T)),
  month = c(sample(5:10, size = 1000, replace = T)),
  year = 2023,
  datum = sample(c('nad83', 'nad27', 'wgs84'), size = 1000, replace = T),
  longitude = runif(1000, min = -120, max = -100), 
  latitude = runif(1000, min = 35, max = 48)
) %>% 
  mutate(col_date = paste(month, day, year, sep = "/"))

```

```{r coordinate parsing and make object spatial in wgs84}

out <- dms2dd(dummy_pts)
out <- coords2sf(out)
```

```{r date parsing}

out <- date_parser(out, coll_date = 'col_date')
```

```{r Acquire geodata}

out1 <- political_grabber(out, 'collectionNo')
out2 <- place_grabber(out1)
out3 <- physical_grabber(out2)

```

```{r gather taxonomic data}

t <- taxize::names_list('species')
t <- c(t, 'agoseris potter')

tout <- spell_check(t)
tres <- lapply(tout$Result, powo_searcher) %>% 
  bind_rows()

notFound(tres)

```

```{r make herbarium label maps, eval = F}

states_sf <- st_read('../geodata/political/political.shp', quiet = T)

dummy_pts_l <- split(dummy_pts, dummy_pts$collectionNo)
out <- lapply(dummy_pts_l, map_maker, states_sf, '../results', collection_col = 'collectionNo')

```