### 2023-06-18 scripts/taxonomy_setup.Rmd
resolved merge via load library

---
title: "Barneby Lives!"
author: "steppe"
date: "9/13/2021"
output: html_document
---

```{r}
library(tidyverse)
library(kewr)
library(WorldFlora)
# WFO.remember()
source('functions.R')
```

# the first two letters of the genus must be correct, 
# the first three letters of the species must be correct

```{r prepare materials for look up tables, eval = F}

cla <- data.table::fread('../taxonomic_data_raw/WFO_Backbone/classification.txt', na.strings = "")

sppLKPtab <- cla[, c('scientificName', 'specificEpithet', 'infraspecificEpithet', 'verbatimTaxonRank')]
sppLKPtab <- filter(sppLKPtab, is.na(verbatimTaxonRank) | verbatimTaxonRank != "f.")

genLKPtab <- unique(cla, by = 'genus')[taxonomicStatus == 'Accepted','genus'] %>% 
  arrange(genus) %>% 
  mutate(Grp = str_extract(genus, '[A-Z][a-z]{1}'))%>% 
  rename(strings = genus)
epiLKPtab <- unique(cla, by = 'specificEpithet')[
  taxonomicStatus == 'Accepted','specificEpithet'] %>% 
  arrange(specificEpithet) %>% 
  mutate(Grp = str_extract(specificEpithet, '[a-z]{3}')) %>% 
  rename(strings = specificEpithet)

write.csv(sppLKPtab, '../taxonomic_data/species_lookup_table.csv', row.names = F)
write.csv(epiLKPtab, '../taxonomic_data/epithet_lookup_table.csv', row.names = F)
write.csv(genLKPtab, '../taxonomic_data/genus_lookup_table.csv', row.names = F)
```

```{r}
lapply(c('Cryptantha glomerata', 'Sphenosciadium capitellatum'), powo_searcher)
```