### 2023-09-03 manuscript/DataSourceTable.Rmd
working on on introduction

---
title: "DataSourcesTable"
author: "steppe"
date: "2023-09-02"
output: html_document
---

```{r echo = F, warning = F, message=F}
library(knitr)
library(kableExtra)
library(gt)
library(tidyverse)
```

```{r Create Table}

local_data <- data.frame(
  Variable = c(
    'County', 'State', 'Ownership', 
    'TRS', 'Place Names', 'Mountains',
    'Elevation', 'Slope', 'Aspect',
    'Geomorphons', 'Surficial Geology', 'Taxonomic Spellings', 
    'Author Abbreviations'
    ),
    Usage = c(
    'Political', 'Political', 'Political',
    'Political', 'Site Name', 'Site Name', 
    'Site Characteristics', 'Site Characteristics', 'Site Characteristics', 
    'Site Characteristics', 'Site Characteristics', 'Spell Checks', 
    'Spell Checks'
    ),
    Source = c(
    'US Census Bureau', 'US Census Bureau', 'US Geological Survey',
    'US Geological Survey', 'US Geological Survey', 'EarthEnv', 
    'Open Topography', 'Open Topography', 'Open Topography',
    'Open Topography', 'World Flora Online', 'US Geological Survey',
    'IPNI'
    ),
  Name = c(
    'Counties', 'States', 'Protected Areas Database', 
    'Public Land Survey System', 'Geographic Names Information System', 'GMBA Mountain Inventory v2',
    'Geomorpho90m - Elevation', 'Geomorpho90m - Slope', 'Geomorpho90m - Aspect',
    'Geomorpho90m - Geomorphons', 'State Geologic Map Compilation', 'World Flora Online',
    'International Plant Names Index'
    ),
  Data_Model = c(
    'Vector', 'Vector', 'Vector',
    'Vector', 'Vector', 'Vector',
    'Raster', 'Raster', 'Raster', 
    'Raster', 'Vector', 'Tabular',
    'Tabular'
    ),
  Subset_Size = c(
    0.073, 0.0, 0.435,
    0.816, 0.081, 0.004,
    4.2, 4.6, 4.1,
    0.455, 0.708, 0.002,
    0.001
  )
)

```