### 2023-09-29 manuscript/labels/skeleton.Rmd
making labels

---
output: pdf_document
params:
  Collection_number: Collection_number
  UNIQUEID: UNIQUEID
geometry: paperheight=4in,paperwidth=4in,margin=0.1in
header-includes:
  - \pagenumbering{gobble}
  - \usepackage{setspace}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dpi = 300) 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r load library}
library(BarnebyLives)
```

```{r load data}
data <- readRDS('../data/processed/data_w_Google_Maps') |>
  sf::st_drop_geometry() |>
  dplyr::filter(Collection_number == params$Collection_number) |>  
  sf::st_drop_geometry()
```

\begin{center}
\textbf{\LARGE Flora of Western North America}
\end{center}
\begin{flushright}
\uppercase{`r data$Family`}
\end{flushright}
\begingroup
\large
`r writer(paste(data$Genus, data$Epithet), italics = TRUE)` `r data$Binomial_authority` `r writer(data$Infraspecific_rank)` `r writer(data$Infraspecies, italics = TRUE)` `r writer(data$Infraspecific_authority)`  
\normalsize
\endgroup

\vspace*{\fill}
\begingroup
`r data$latitude_dd` `r data$longitude_dd` (`r data$Datum`). `r data$Gen`. `r writer(data$Site)`

`r writer(data$Habitat)` `r data$physical_environ`  
Veg.: `r species_font(data$Vegetation)` `r associates_writer(data$Associates)`
`r writer(data$Notes)`  
\endgroup
\begingroup
\begin{spacing}{0.1}
{\scriptsize `r data$Directions_BL`}
\end{spacing}
\endgroup

\vspace*{\fill}
\begingroup
`r collection_writer(data)`. `r writer_fide(data)`
\endgroup

\begin{center}
{\footnotesize Collected under the auspices of the Bureau of Land Management}
\end{center}