### 2025-04-24 manuscript/CoverLetter.rmd

---
output:
  pdf_document: default
  word_document: default
---

\begin{flushleft}
Brianna Gross \newline
Editor in Chief \newline
Applications in Plant Sciences \newline
\end{flushleft}

\hfill April 24^th^, 2025

Dear Dr. Gross, 
\vspace{3em}

I am writing to propose an original software article titled *"BarnebyLives: an R package to create herbarium specimen labels and clean spreadsheets"*

I have no conflicts of interest to disclose.