### 2025-05-01 SubsetRecords2CBG.rmd

---
author: "Reed Benkendorf"
date: "2025-04-29"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
p <- './manuscript/data'

dat_sub <- read.csv(file.path(p, 'test_data.csv')) |>
  select(Collection_number, CBG_ACC_BOOL) |>
  mutate(CBG_ACC_BOOL = if_else(is.na(CBG_ACC_BOOL), TRUE, FALSE)) |>
  filter(CBG_ACC_BOOL == TRUE)

read.csv( file.path(p, 'processed', 'cleaned_data.csv') ) |>
  select(-CBG_ACC_BOOL) |>
  right_join(dat_sub)
```