### 2025-05-01 SubsetRecords2CBG.rmd
data writeout for CHIC - ready, do a final manual review

---
author: "Reed Benkendorf"
date: "2025-04-29"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
p <- './manuscript/data/processed'

dat_sub <- read.csv(file.path(p, 'cleaned_data.csv')) |>
  mutate(CBG_ACC_BOOL = if_else(is.na(CBG_ACC_BOOL), TRUE, FALSE)) 
dat_sub <- data.frame( apply(dat_sub, 2, as.character) )

dat_manual <- read.csv(file.path(p, 'cleaned_data-problemchildren.csv')) |>
  mutate(CBG_ACC_BOOL = if_else(is.na(CBG_ACC_BOOL), TRUE, FALSE)) 
dat_manual <- data.frame( apply(dat_manual, 2, as.character) )

dat_out <- filter(dat_sub, ! Collection_number %in% dat_manual$Collection_number) |>
  bind_rows( dat_manual) |>
  filter(CBG_ACC_BOOL == 'TRUE') |>
  arrange(Collection_number)

write.csv(dat_out, file.path(p, 'CHIC-RCB-2025.csv'))
```