### 2025-08-09 LabelStyles/skeleton-maps.Rmd

---
output: pdf_document
params:
  Collection_number: Collection_number
  imagepath: imagepath
geometry: paperheight=4in,paperwidth=4in,margin=0.1in
header-includes:
  - \pagenumbering{gobble}
  - \usepackage{setspace}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dpi = 300) 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r load library}
library(BarnebyLives)
```

```{r load data}
record <- collection_examples |>
  dplyr::filter(Collection_number == params$Collection_number) |>  
  sf::st_drop_geometry()

imagepath <- params$imagepath
```

\begin{center}
\textbf{\Large `r writer(record[['Project_Name']])`}
\end{center}
\begin{flushright}
\uppercase{`r record[['Family']]`}
\end{flushright}

\noindent
\begin{minipage}[t]{0.78\textwidth}\vspace*{0pt}

\begingroup
\large
`r writer(paste(record[['Genus']], record[['Epithet']]), italics = TRUE)` `r writer(record[['Binomial_authority']])` `r writer(record[['Infrarank']])` `r writer(record[['Infraspecies']], italics = TRUE)` `r writer(record[['Infraspecific_authority']])`  
`r writer_fide(record)`
\normalsize
\endgroup

\vspace*{\fill}

\begingroup
`r record[['Gen']]`. `r writer(record[['Site']])` `r record[['latitude_dd']]` `r record[['longitude_dd']]` (`r record[['Datum']]` `r writer(record[['Coordinate_Uncertainty']])`).

`r writer(record[['Habitat']])`. `r record[['physical_environ']]`  
Veg.: `r species_font(record[['Vegetation']])` `r associates_writer(record[['Associates']])`
`r writer(record[['Notes']])`  
\endgroup

\vspace*{\fill}
\begingroup
`r collection_writer(record)`
\endgroup
\end{minipage}%
\hfill
\begin{minipage}[c]{0.18\textwidth}\vspace*{0pt}
\hspace*{4mm}% left margin
\centering
\includegraphics[width=\linewidth,height=0.66\textheight,keepaspectratio]{`r imagepath`}
\end{minipage}

convert -density 300 Molecular_nets-crop.pdf -quality 100 \
        -gravity West -background white -splice 100x0 \
        \( -background transparent -fill black -pointsize 12 \
           label:"Time within Season" -rotate -90 \) \
        -geometry +0+250 -composite Molecular_nets-crop-labelled.png