### 2025-10-12 valleys2/pull_hu10_for_valleys_project_oct10_2025.qmd
Reworked Valleys dataset for BL

---
title: "Pull hu10"
format: html
editor: visual
---

```{r}
library(sf)
library(tidyverse)
library(rmapshaper)
getwd()
```

```{r, eval = F}
p <- '/media/steppe/hdd/Geospatial_data/NHDPlusNationalData/NHDPlusV21_National_Seamless_Flattened_Lower48.gdb'
st_layers(p)
hu10 <- st_read(p, layer = 'HUC12') |>
  st_make_valid()

domain <- st_read('/media/steppe/hdd/Barneby_Lives-dev/ValleysProject/data/domain.gpkg') |>
  st_transform(st_crs(hu10))

hu10 <- hu10[ unlist(st_intersects(domain, hu10)),] |>
  st_make_valid()

mountains <- st_read('/media/steppe/hdd/Barneby_Lives-dev/geodata/mountains') |>
  st_make_valid() |>
  st_transform(st_crs(hu10)) |>
  st_union()

mountains <- sf::st_set_precision(mountains, precision = 1e3) |>
  st_make_valid()
hu10 <- sf::st_set_precision(hu10, precision = 1e3)|>
  st_make_valid()

hu10 <- st_difference(hu10, mountains)
hu10 <- st_make_valid(hu10)

polys <- hu10[complete.cases(
  sf::st_coordinates(sf::st_centroid(sf::st_geometry(hu10)))
), ]

polys <- polys %>%
  sf::st_collection_extract("POLYGON")

polys <- polys %>%
    st_make_valid() %>%                     # Step 1: Repair invalid geometries
    st_cast("MULTIPOLYGON", warn = FALSE) %>%  # Step 2: Ensure MULTIPOLYGON type
    st_cast("POLYGON", warn = FALSE) %>%       # Step 3: Flatten to POLYGONs
    filter(!st_is_empty(.))           # Final: Drop empty geometries

st_write(polys, '/media/steppe/hdd/Barneby_Lives-dev/ValleysProject/data/Main_huc12_polygons.gpkg', append = F)
#hu10 <- rmapshaper::ms_simplify(hu10, keep = 0.05, keep_shapes = TRUE)
```

```{r}
polys <- st_read(
  '/media/steppe/hdd/Barneby_Lives-dev/ValleysProject/data/Main_huc12_polygons.gpkg') |>
  sf::st_make_valid()

polys_simp <- rmapshaper::ms_simplify(polys, keep = 0.1, keep_shapes  = TRUE)
polys_simp <- polys_simp[!st_is_empty(polys_simp),]

st_write(
  polys_simp, 
  '/media/steppe/hdd/Barneby_Lives-dev/ValleysProject/data/Simplified_huc12_polygons.gpkg', 
  append = F)
```