### 2025-10-25 LabelStyles/skeleton-default.Rmd
label changes - check diff on vignettes

---
output: pdf_document
params:
  Collection_number: NULL
geometry: paperheight=4in,paperwidth=4in,margin=0.1in
header-includes:
  - \pagenumbering{gobble}
  - \usepackage{setspace}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi = 300) 
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r load data, eval = T, echo= F}
library(BarnebyLives)

local <- file.path('~', 'Documents', 'assoRted', 'Barneby_Lives_dev', 'LabelStyles')

record <- read.csv(file.path(local, 'SoS-ExampleCollection.csv')) |>
  sf::st_drop_geometry() |>
  data.frame() |>
  dplyr::filter(Collection_number == params$Collection_number)

# escape character columns to ensure characters which interfere with latex
# rendering are escaped. 
char_cols <- sapply(record, is.character)
char_col_names <- names(record)[char_cols]
record[char_col_names] <- lapply(record[char_col_names], escape_latex)
```

\begin{center}
\textbf{\Large `r writer(record[['Project_name']])`}
\end{center}
\begin{flushright}
\uppercase{`r record[['Family']]`}
\end{flushright}
\begingroup
\large

`r writer(paste(record[['Genus']], record[['Epithet']]), italics = TRUE)` `r writer(record[['Binomial_authority']])` `r writer(record[['Infrarank']])` `r writer(record[['Infraspecies']], italics = TRUE)` `r writer(record[['Infraspecific_authority']])`\
\endgroup

\vspace{\fill}
\begingroup
\normalsize

`r record[['Gen']]`. `r writer(record[['Site']])` `r record[['latitude_dd']]` `r record[['longitude_dd']]` (`r record[['Datum']]` `r writer(record[['Coordinate_uncertainty']])`).

`r writer(record[['Habitat']])`. `r record[['physical_environ']]`

Vegetation: `r species_font(record[['Vegetation']])`

`r associates_writer(record[['Associates']], full = TRUE)` \endgroup

\vspace{\fill}
\begingroup

`r collection_writer(record, date=FALSE)`

\begin{flushright}
`r record[['Date_digital_text']]`
\end{flushright}
\endgroup
\begingroup
\small
`r writer(record[['Notes']])`
\endgroup