### 2025-11-20 LabelStyles/SoS-skeleton-Det.Rmd
resubmitted version

---
output: pdf_document
params:
  Collection_number: Collection_number
geometry: paperheight=4in,paperwidth=4in,margin=0.1in
header-includes:
  - \pagenumbering{gobble}
  - \usepackage{setspace}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dpi = 300) 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r load library}
library(BarnebyLives)
```

```{r load data}
record <- read.csv('SoS-ExampleCollection.csv') |>
  dplyr::mutate(
    Collection_number = as.numeric(Collection_number), 
    Coordinate_uncertainty = '+/- 5m') |>
  dplyr::filter(Collection_number == params$Collection_number) |>
  dplyr::mutate(Project_name = 'SOS Template')
```

\begin{center}
\textbf{\LARGE `r record[['Project_name']]`}
\end{center}
\vspace{-\baselineskip} 
\uppercase{`r record[['Family']]`}  
\vspace{-\baselineskip} 
\begingroup
\large
`r writer(paste(record[['Genus']], record[['Epithet']]), italics = TRUE)` `r writer(record[['Binomial_authority']])` `r writer(record[['Infrarank']])` `r writer(record[['Infraspecies']], italics = TRUE)` `r writer(record[['Infraspecific_authority']])`  
\normalsize 
`r writer_fide(record)`
\endgroup

\vspace{\fill}
\begingroup
\small
`r record[['Gen']]`. `r writer(record[['Site']])` `r record[['latitude_dd']]` `r record[['longitude_dd']]` (`r record[['Datum']]` `r writer(record[['Coordinate_uncertainty']])`).

`r writer(record$Habitat)`. `r record[['physical_environ']]`

Veg.: `r species_font(record[['Vegetation']])` `r associates_writer(record[['Associates']])`
`r writer(record[['Notes']])`  
\endgroup

\par\vspace{\fill}

\begingroup
\small
`r collection_writer(record)`
\endgroup

\vspace{-\baselineskip} 
\begin{center}
{\footnotesize Collected under the auspices of the Bureau of Land Management}
\end{center}