### 2025-11-20 LabelStyles/creating_example_labels.Rmd

---
title: "Printing herbarium labels and exporting a digital copy of data"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Printing herbarium labels and exporting a digital copy of data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
description: >
  Rendering herbarium labels
---

```{r load libraries and import data}
# remotes::install_github('sagesteppe/BarnebyLives') 
# OR devtools::install_github('sagesteppe/BarnebyLives')
library(BarnebyLives)
library(dplyr)
library(glue)
data('collection_examples')
```

## creating a driver script to feed the skeleton data

```{r}
p <- file.path(path.expand('~'), 'Documents', 'assoRted', 'Barneby_Lives_dev', 'LabelStyles')
driver <- function(x, name){
  purrr::walk(
    .x = 2823,
  ~ rmarkdown::render(
    input = x,                      # template to populate
    output_file = file.path(p, glue("{name}{.x}.pdf")),# write one PDF per record
    params = list(Collection_number = {.x})      # pass record to skeleton
    )
  )
}
```

### BUILT IN TEMPLATES

```{r, eval = F}
driver("skeleton-default.Rmd", name = 'default')

driver("skeleton-research.Rmd", name = 'research')
driver("skeleton-modern.Rmd", name = 'modern')
driver("skeleton-SOS.Rmd", name = 'SOS')
driver("skeleton-customfont.Rmd", name = 'customfont')
```

### LABELS WITH GRAPHICS

```{r}
#driver("skeleton-maps.Rmd", name = 'map')
#driver("skeleton-default.Rmd", name = 'default-')
```

## customizing a label template

```{r customizing label templates, eval = F}
driver("skeleton-SOS.Rmd", name = 'default-SOS-')
driver("SoS-skeleton-noDir.Rmd", name = 'noDir-SOS-')
driver("SoS-skeleton-Spaced.Rmd", name = 'Spaced-SOS-')
driver("SoS-skeleton-large.Rmd", name = 'large-SOS-')
driver("SoS-skeleton-familyLeft.Rmd", name = 'familyLeft-SOS-')
driver("SoS-skeleton-Det.Rmd", name = 'Det-SOS-')
```

## customizing label fonts

(completed)

```{r, eval = F}
driver("skeleton-customfont.Rmd", name = 'custom-font')
driver("skeleton-customfont-twofonts.Rmd", name = 'custom-twofonts')
driver("skeleton-customfont-downloaded.Rmd", name = 'custom-font-downloaded')
```