### 2025-11-20 LabelStyles/skeleton-customfont-downloaded.Rmd

---
output: 
  pdf_document:
    latex_engine: xelatex # or lualatex
    keep_tex: true
params:
  Collection_number: Collection_number
geometry: paperheight=4in,paperwidth=4in,margin=0.1in
header-includes:
  - \pagenumbering{gobble}
  - \usepackage{setspace}
  - \usepackage{fontspec}
  - |
    \setmainfont[
      Path = /home/sagesteppe/Documents/assoRted/Barneby_Lives_dev/LabelStyles/Nunito/static/,
      UprightFont = *-Regular,
      ItalicFont = *-Italic,
      BoldFont = *-Bold,
      BoldItalicFont = *-BoldItalic
    ]{Nunito}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dpi = 300) 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r load library}
library(BarnebyLives)
```

```{r load data}
local <- file.path('~', 'Documents', 'assoRted', 'Barneby_Lives_dev', 'LabelStyles')

record <- read.csv(file.path(local, 'SoS-ExampleCollection.csv')) |>
  dplyr::filter(Collection_number == params$Collection_number)  |>
  dplyr::mutate(Project_name = 'Custom font downloaded')
```

\begin{center}
\textbf{\Large `r writer(record[['Project_name']])`}
\end{center}
\begin{flushright}
\uppercase{`r record[['Family']]`}
\end{flushright}
\begingroup
\large
`r writer(paste(record[['Genus']], record[['Epithet']]), italics = TRUE)` `r writer(record[['Binomial_authority']])` `r writer(record[['Infrarank']])` `r writer(record[['Infraspecies']], italics = TRUE)` `r writer(record[['Infraspecific_authority']])`  
\normalsize
\endgroup

\vspace{\fill}
\begingroup
`r record[['Gen']]`. `r writer(record[['Site']])` `r record[['latitude_dd']]` `r record[['longitude_dd']]` (`r record[['Datum']]` `r writer(record[['Coordinate_uncertainty']])`).

`r writer(record[['Habitat']])`. `r record[['physical_environ']]` \par
Vegetation: `r species_font(record[['Vegetation']])` \par
`r associates_writer(record[['Associates']])`
`r writer(record[['Notes']])`  
\endgroup

\vspace{\fill}
\begingroup
`r collection_writer(record)` \par
`r writer_fide(record)`
\endgroup