### 2025-11-20 manuscript/Appendices.Rmd

---
output:
  pdf_document: default
---
# Appendix 1: System Information for Benchmarking

The following system details are provided to support reproducibility of the performance measurements reported in this work.
All benchmarks were conducted on a local machine with the specifications shown below. 
Information was collected via standard system commands at time of receiving reviewer comments.

## Operating System

```{sh, eval = F}
lsb_release -a
```

Distributor ID:	Ubuntu  
Description:	Ubuntu 22.04.5 LTS  
Release:	22.04  
Codename:	jammy

## CPU

```{sh, eval = F}
lscpu
```

| Field               | Value                                          |
| ------------------- | ---------------------------------------------- |
| Architecture        | x86_64                                         |
| CPU op-mode(s)      | 32-bit, 64-bit                                 |
| Address sizes       | 39 bits physical, 48 bits virtual              |
| Byte Order          | Little Endian                                  |
| CPU(s)              | 16                                             |
| On-line CPU(s) list | 0-15                                           |
| Vendor ID           | GenuineIntel                                   |
| Model name          | 11th Gen Intel(R) Core(TM) i7-11700K @ 3.60GHz |
| CPU family          | 6                                              |
| Model               | 167                                            |
| Thread(s) per core  | 2                                              |
| Core(s) per socket  | 8                                              |
| Socket(s)           | 1                                              |
| Stepping            | 1                                              |
| CPU max MHz         | 5100.0000                                      |
| CPU min MHz         | 800.0000                                       |
| BogoMIPS            | 7200.00                                        |

\newpage

## Memory

```{sh, eval = F}
free -h
```

| Type | Total | Used  | Free  | Shared | Buff/Cache | Available |
| ---- | ----- | ----- | ----- | ------ | ---------- | --------- |
| Mem  | 125Gi | 3.1Gi | 112Gi | 569Mi  | 9.7Gi      | 120Gi     |
| Swap | 2.0Gi | 0B    | 2.0Gi | —      | —          | —         |

## Internet Speed

```{sh, eval = F}
speedtest-cli
```

Retrieving speedtest.net configuration...  
Testing from Astound Broadband (168.91.196.6)...  
Retrieving speedtest.net server list...  
Selecting best server based on ping...  
Hosted by Sangoma (Chicago, IL) [17.63 km]: 18.388 ms  
Testing download speed..........................................................  
Download: 21.58 Mbit/s  
Testing upload speed............................................................  
Upload: 27.43 Mbit/s

## Filesystem Usage

```{sh, eval = F}
df -h
```

| Filesystem     | Size | Used | Avail | Use% | Mounted on                |
| -------------- | ---- | ---- | ----- | ---- | ------------------------- |
| tmpfs          | 13G  | 2.6M | 13G   | 1%   | /run                      |
| /dev/nvme0n1p2 | 468G | 242G | 203G  | 55%  | /                         |
| tmpfs          | 63G  | 14M  | 63G   | 1%   | /dev/shm                  |
| tmpfs          | 5.0M | 4.0K | 5.0M  | 1%   | /run/lock                 |
| efivarfs       | 192K | 88K  | 100K  | 47%  | /sys/firmware/efi/efivars |
| /dev/nvme0n1p1 | 511M | 6.1M | 505M  | 2%   | /boot/efi                 |
| tmpfs          | 13G  | 132K | 13G   | 1%   | /run/user/1000            |
| /dev/sda1      | 3.6T | 3.2T | 295G  | 92%  | /media/steppe/hdd         |

## I/O Performance

```{sh, eval = F}
sudo hdparm -Tt /dev/nvme0n1p2
```

Timing cached reads:   44096 MB in  2.00 seconds = 22095.91 MB/sec  
Timing buffered disk reads: 4470 MB in  3.00 seconds = 1489.83 MB/sec

All files were stored on the following drive.

```{sh, eval = F}
sudo hdparm -Tt /dev/sda1
```

Timing cached reads:   45870 MB in  2.00 seconds = 22986.94 MB/sec  
Timing buffered disk reads: 514 MB in  3.01 seconds = 170.57 MB/sec

\newpage

# Appendix 2: Label Templates

\begin{center}
\includegraphics[width=420pt]{./labels/TemplateExamples.png}
\end{center}